---
title: "Booster Model Formulae"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{booster_model_formulae}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Age groups and Vaccination Statuses

The model includes 17 age groups (indexed by $a$) and 8 vaccination statuses (indexed by $v$), where:

- $v = 1$: Unvaccinated
- $v = 2$: First dose received
- $v = 3$: Second dose received
- $v = 4, 5$: Waned immunity states
- $v = 6$: First booster received
- $v = 7, 8$: Further waned booster states

# Model Compartments

The model consists of the following compartments for each age group $a$ and vaccination status $v$:

- $S_{a,v}$: Susceptible individuals
- $E1_{a,v}$, $E2_{a,v}$: Exposed individuals (latent infection stages)
- $I^{\text{Mild}}_{a,v}$: Mild infections (non-hospitalized)
- $I1_{a,v}$, $I2_{a,v}$: Severe infections (pre-hospitalization)
- $R1_{a,v}$, $R2_{a,v}$: Recovered individuals
- $D_{a,v}$: Deaths

Hospital and ICU compartments (detailed in Section 4):

- Hospitalised compartments: $H^{\text{Hosp,Get,Live}}_{a,v}$, $H^{\text{Hosp,Get,Die}}_{a,v}$, $H^{\text{Hosp,NotGet,Live}}_{a,v}$, $H^{\text{Hosp,NotGet,Die}}_{a,v}$
- ICU compartments: $H^{\text{ICU,Get,Live}}_{a,v}$, $H^{\text{ICU,Get,Die}}_{a,v}$, $H^{\text{ICU,NotGet,Live}}_{a,v}$, $H^{\text{ICU,NotGet,Die}}_{a,v}$
- Recovery (from ICU) compartments: $H1^{\text{Rec}}_{a,v}$, $H2^{\text{Rec}}_{a,v}$

Please note that compartment values vary over time, but this notation is not included for brevity, i.e. $S_{a,v}$ is truly $S_{a,v}(t)$.

# Core Epidemiological Dynamics

## Susceptible Compartment

$$\frac{dS_{a,1}}{dt} = \gamma_R(t) R2_{a,1} - \lambda_a \xi^{\text{inf}}_1(t) S_{a,1} - \gamma^{\text{vac}}_1(t) S_{a,1} + V^S_{a,1}$$

$$\frac{dS_{a,v}}{dt} = \gamma_R(t) R2_{a,v} - \lambda_a \xi^{\text{inf}}_v(t) S_{a,v} - \gamma^{\text{vac}}_v(t) S_{a,v} + V^S_{a,v} + \gamma^{\text{vac}}_{v-1}(t) S_{a,v-1}$$

for $v = 2, \ldots, 8$

where:

- $\lambda_a$ is the force of infection for age group $a$
- $\xi^{\text{inf}}_v(t)$ is the vaccine efficacy against infection for vaccination status $v$
- $\gamma^{\text{vac}}_v(t)$ is the rate of vaccine waning for status $v$
- $\gamma_R(t)$ is the rate of loss of natural immunity
- $V^S_{a,v}$ represents vaccination transitions

## Exposed Compartments

$$\frac{dE1_{a,1}}{dt} = \lambda_a \xi^{\text{inf}}_1(t) S_{a,1} - \gamma_E E1_{a,1} - \gamma^{\text{vac}}_1(t) E1_{a,1} + V^{E1}_{a,1}$$

$$\frac{dE1_{a,v}}{dt} = \lambda_a \xi^{\text{inf}}_v(t) S_{a,v} - \gamma_E E1_{a,v} - \gamma^{\text{vac}}_v(t) E1_{a,v} + V^{E1}_{a,v} + \gamma^{\text{vac}}_{v-1}(t) E1_{a,v-1}$$

for $v = 2, \ldots, 8$

$$\frac{dE2_{a,1}}{dt} = \gamma_E E1_{a,1} - \gamma_E E2_{a,1} - \gamma^{\text{vac}}_1(t) E2_{a,1} + V^{E2}_{a,1}$$

$$\frac{dE2_{a,v}}{dt} = \gamma_E E1_{a,v} - \gamma_E E2_{a,v} - \gamma^{\text{vac}}_v(t) E2_{a,v} + V^{E2}_{a,v} + \gamma^{\text{vac}}_{v-1}(t) E2_{a,v-1}$$

for $v = 2, \ldots, 8$

where $\gamma_E$ is the rate of progression through latent infection.

## Mild Infections

$$\frac{dI^{\text{Mild}}_{a,1}}{dt} = \gamma_E E2_{a,1} (1 - p^{\text{hosp}}_{a,1}(t)) - \gamma^{\text{Mild}} I^{\text{Mild}}_{a,1} - \gamma^{\text{vac}}_1(t) I^{\text{Mild}}_{a,1}$$

$$\frac{dI^{\text{Mild}}_{a,v}}{dt} = \gamma_E E2_{a,v} (1 - p^{\text{hosp}}_{a,v}(t)) - \gamma^{\text{Mild}} I^{\text{Mild}}_{a,v} - \gamma^{\text{vac}}_v(t) I^{\text{Mild}}_{a,v} + \gamma^{\text{vac}}_{v-1}(t) I^{\text{Mild}}_{a,v-1}$$

for $v = 2, \ldots, 8$

where:

- $p^{\text{hosp}}_{a,v}(t) = p^{\text{hosp}}_a \cdot \xi^{\text{dis}}_v(t) \cdot m^{\text{hosp}}(t)$ is the time-varying hospitalization probability
- $\gamma^{\text{Mild}}$ is the recovery rate from mild infection

## Severe Infections (Pre-hospitalization)

$$\frac{dI1_{a,1}}{dt} = \gamma_E E2_{a,1} p^{\text{hosp}}_{a,1}(t) - \gamma^{\text{Case}} I1_{a,1} - \gamma^{\text{vac}}_1(t) I1_{a,1}$$

$$\frac{dI1_{a,v}}{dt} = \gamma_E E2_{a,v} p^{\text{hosp}}_{a,v}(t) - \gamma^{\text{Case}} I1_{a,v} - \gamma^{\text{vac}}_v(t) I1_{a,v} + \gamma^{\text{vac}}_{v-1}(t) I1_{a,v-1}$$

$$\frac{dI2_{a,1}}{dt} = \gamma^{\text{Case}} I1_{a,1} - \gamma^{\text{Case}} I2_{a,1} - \gamma^{\text{vac}}_1(t) I2_{a,1}$$

$$\frac{dI2_{a,v}}{dt} = \gamma^{\text{Case}} I1_{a,v} - \gamma^{\text{Case}} I2_{a,v} - \gamma^{\text{vac}}_v(t) I2_{a,v} + \gamma^{\text{vac}}_{v-1}(t) I2_{a,v-1}$$

for $v = 2, \ldots, 8$

## Recovered Compartments

$$\begin{align}
\frac{dR1_{a,1}}{dt} &= \gamma^{\text{rec}} H2^{\text{Rec}}_{a,1} + \gamma^{\text{Mild}} I^{\text{Mild}}_{a,1} + \gamma^{\text{Hosp,surv}}(t) H2^{\text{Hosp,Get,Live}}_{a,1} \\
&\quad + \gamma^{\text{Hosp,surv}} H2^{\text{Hosp,NotGet,Live}}_{a,1} + \gamma^{\text{ICU,surv}} H2^{\text{ICU,NotGet,Live}}_{a,1} \\
&\quad - \gamma_R(t) R1_{a,1} - \gamma^{\text{vac}}_1(t) R1_{a,1} + V^{R1}_{a,1}
\end{align}$$

$$\begin{align}
\frac{dR1_{a,v}}{dt} &= \gamma^{\text{rec}} H2^{\text{Rec}}_{a,v} + \gamma^{\text{Mild}} I^{\text{Mild}}_{a,v} + \gamma^{\text{Hosp,surv}}(t) H2^{\text{Hosp,Get,Live}}_{a,v} \\
&\quad + \gamma^{\text{Hosp,surv}} H2^{\text{Hosp,NotGet,Live}}_{a,v} + \gamma^{\text{ICU,surv}} H2^{\text{ICU,NotGet,Live}}_{a,v} \\
&\quad - \gamma_R(t) R1_{a,v} - \gamma^{\text{vac}}_v(t) R1_{a,v} + V^{R1}_{a,v} + \gamma^{\text{vac}}_{v-1}(t) R1_{a,v-1}
\end{align}$$

for $v = 2, \ldots, 8$

$$\frac{dR2_{a,1}}{dt} = \gamma_R(t) R1_{a,1} - \gamma_R(t) R2_{a,1} - \gamma^{\text{vac}}_1(t) R2_{a,1} + V^{R2}_{a,1}$$

$$\frac{dR2_{a,v}}{dt} = \gamma_R(t) R1_{a,v} - \gamma_R(t) R2_{a,v} - \gamma^{\text{vac}}_v(t) R2_{a,v} + V^{R2}_{a,v} + \gamma^{\text{vac}}_{v-1}(t) R2_{a,v-1}$$

for $v = 2, \ldots, 8$

# Hospital and ICU Dynamics

For individuals who do not require ICU, get treatment and survive:

$$\begin{align}
\frac{dH1^{\text{Hosp,Get,Live}}_{a,1}}{dt} &= \gamma^{\text{Case}} I2_{a,1} (1 - p^{\text{sev}}_a(t)) p^{\text{hosp}} (1 - p^{\text{death,Hosp,treat}}_a) \\
&\quad - \gamma^{\text{Hosp,surv}}(t) H1^{\text{Hosp,Get,Live}}_{a,1} - \gamma^{\text{vac}}_1(t) H1^{\text{Hosp,Get,Live}}_{a,1}
\end{align}$$

$$\begin{align}
\frac{dH1^{\text{Hosp,Get,Live}}_{a,v}}{dt} &= \gamma^{\text{Case}} I2_{a,v} (1 - p^{\text{sev}}_a(t)) p^{\text{hosp}} (1 - p^{\text{death,Hosp,treat}}_a) \\
&\quad - \gamma^{\text{Hosp,surv}}(t) H1^{\text{Hosp,Get,Live}}_{a,v} - \gamma^{\text{vac}}_v(t) H1^{\text{Hosp,Get,Live}}_{a,v} \\
&\quad + \gamma^{\text{vac}}_{v-1}(t) H1^{\text{Hosp,Get,Live}}_{a,v-1}
\end{align}$$

for $v = 2, \ldots, 8$

$$\frac{dH2^{\text{Hosp,Get,Live}}_{a,1}}{dt} = \gamma^{\text{Hosp,surv}}(t) H1^{\text{Hosp,Get,Live}}_{a,1} - \gamma^{\text{Hosp,surv}}(t) H2^{\text{Hosp,Get,Live}}_{a,1} - \gamma^{\text{vac}}_1(t) H2^{\text{Hosp,Get,Live}}_{a,1}$$

$$\begin{align}
\frac{dH2^{\text{Hosp,Get,Live}}_{a,v}}{dt} &= \gamma^{\text{Hosp,surv}}(t) H1^{\text{Hosp,Get,Live}}_{a,v} - \gamma^{\text{Hosp,surv}}(t) H2^{\text{Hosp,Get,Live}}_{a,v} \\
&\quad - \gamma^{\text{vac}}_v(t) H2^{\text{Hosp,Get,Live}}_{a,v} + \gamma^{\text{vac}}_{v-1}(t) H2^{\text{Hosp,Get,Live}}_{a,v-1}
\end{align}$$

for $v = 2, \ldots, 8$

Similar equations exist for:

- $H1^{\text{Hosp,Get,Die}}_{a,v}$, $H2^{\text{Hosp,Get,Die}}_{a,v}$: Get treatment, die
- $H1^{\text{Hosp,NotGet,Live}}_{a,v}$, $H2^{\text{Hosp,NotGet,Live}}_{a,v}$: Don't get treatment, survive
- $H1^{\text{Hosp,NotGet,Die}}_{a,v}$, $H2^{\text{Hosp,NotGet,Die}}_{a,v}$: Don't get treatment, die

## Mechanical Ventilation Compartments

For individuals who get mechanical ventilation and survive:

$$\begin{align}
\frac{dH1^{\text{ICU,Get,Live}}_{a,1}}{dt} &= \gamma^{\text{Case}} I2_{a,1} p^{\text{sev}}_a(t) p^{\text{ICU}} (1 - p^{\text{death,ICU,treat}}_a) \\
&\quad - \gamma^{\text{ICU,surv}}(t) H1^{\text{ICU,Get,Live}}_{a,1} - \gamma^{\text{vac}}_1(t) H1^{\text{ICU,Get,Live}}_{a,1}
\end{align}$$

$$\begin{align}
\frac{dH1^{\text{ICU,Get,Live}}_{a,v}}{dt} &= \gamma^{\text{Case}} I2_{a,v} p^{\text{sev}}_a(t) p^{\text{ICU}} (1 - p^{\text{death,ICU,treat}}_a) \\
&\quad - \gamma^{\text{ICU,surv}}(t) H1^{\text{ICU,Get,Live}}_{a,v} - \gamma^{\text{vac}}_v(t) H1^{\text{ICU,Get,Live}}_{a,v} \\
&\quad + \gamma^{\text{vac}}_{v-1}(t) H1^{\text{ICU,Get,Live}}_{a,v-1}
\end{align}$$

for $v = 2, \ldots, 8$

Similar structures exist for all mechanical ventilation compartments.

## Recovery from ICU

$$\frac{dH1^{\text{Rec}}_{a,1}}{dt} = \gamma^{\text{ICU,surv}}(t) H2^{\text{ICU,Get,Live}}_{a,1} - \gamma^{\text{rec}} H1^{\text{Rec}}_{a,1} - \gamma^{\text{vac}}_1(t) H1^{\text{Rec}}_{a,1}$$

$$\begin{align}
\frac{dH1^{\text{Rec}}_{a,v}}{dt} &= \gamma^{\text{ICU,surv}}(t) H2^{\text{ICU,Get,Live}}_{a,v} - \gamma^{\text{rec}} H1^{\text{Rec}}_{a,v} \\
&\quad - \gamma^{\text{vac}}_v(t) H1^{\text{Rec}}_{a,v} + \gamma^{\text{vac}}_{v-1}(t) H1^{\text{Rec}}_{a,v-1}
\end{align}$$

for $v = 2, \ldots, 8$

$$\frac{dH2^{\text{Rec}}_{a,1}}{dt} = \gamma^{\text{rec}} H1^{\text{Rec}}_{a,1} - \gamma^{\text{rec}} H2^{\text{Rec}}_{a,1} - \gamma^{\text{vac}}_1(t) H2^{\text{Rec}}_{a,1}$$

$$\begin{align}
\frac{dH2^{\text{Rec}}_{a,v}}{dt} &= \gamma^{\text{rec}} H1^{\text{Rec}}_{a,v} - \gamma^{\text{rec}} H2^{\text{Rec}}_{a,v} \\
&\quad - \gamma^{\text{vac}}_v(t) H2^{\text{Rec}}_{a,v} + \gamma^{\text{vac}}_{v-1}(t) H2^{\text{Rec}}_{a,v-1}
\end{align}$$

for $v = 2, \ldots, 8$

## Deaths

$$\begin{align}
\frac{dD_{a,v}}{dt} &= \gamma^{\text{Hosp,die}}(t) H2^{\text{Hosp,Get,Die}}_{a,v} + \gamma^{\text{Hosp,die}} H2^{\text{Hosp,NotGet,Die}}_{a,v} \\
&\quad + \gamma^{\text{ICU,die}}(t) H2^{\text{ICU,Get,Die}}_{a,v} + \gamma^{\text{ICU,die}} H2^{\text{ICU,NotGet,Die}}_{a,v}
\end{align}$$

# Vaccination Dynamics

## Vaccination Transitions

The vaccination transitions $V^X_{a,v}$ for compartment $X$ are defined as:

$$\begin{align}
V^X_{a,1} &= -r^{\text{prim,1}}_a X_{a,1} \\
V^X_{a,2} &= r^{\text{prim,1}}_a X_{a,1} - r^{\text{prim,2}} X_{a,2} \\
V^X_{a,3} &= r^{\text{prim,2}} X_{a,2} - r^{\text{boost,1}} X_{a,3} \\
V^X_{a,4} &= -r^{\text{boost,1}} X_{a,4} \\
V^X_{a,5} &= -r^{\text{boost,1}} X_{a,5} \\
V^X_{a,6} &= r^{\text{boost,1}} \sum_{k=3}^5 X_{a,k} + r^{\text{boost,2}}_a \sum_{k=7}^8 X_{a,k} \\
V^X_{a,7} &= -r^{\text{boost,2}}_a X_{a,7} \\
V^X_{a,8} &= -r^{\text{boost,2}}_a X_{a,8}
\end{align}$$

where:

- $r^{\text{prim,1}}_a$ is the first dose rate for age group $a$
- $r^{\text{prim,2}}$ is the second dose rate
- $r^{\text{boost,1}}$ is the first booster rate
- $r^{\text{boost,2}}_a$ is the second booster rate for age group $a$

# Force of Infection

The force of infection for age group $a$ is:

$$\lambda_a = \beta(t) \sum_{k=1}^{17} M_{a,k} \rho_k \sum_{v=1}^8 \left( I^{\text{Mild}}_{k,v} + I1_{k,v} + I2_{k,v} \right) \psi_v$$

where:

- $\beta(t)$ is the time-varying transmission rate
- $M_{a,k}$ is the contact matrix between age groups $a$ and $k$
- $\rho_k$ is the relative infectiousness of age group $k$
- $\psi_v$ is the relative infectiousness of vaccination status $v$

# Hospital Capacity Constraints

## Treatment Availability

The probability of receiving (non-ICU) treatment is:

$$p^{\text{hosp}} = \begin{cases}
1 & \text{if } N^{\text{hosp,req}} \leq C^{\text{hosp}} - O^{\text{hosp}} \\
\frac{C^{\text{hosp}} - O^{\text{hosp}}}{N^{\text{hosp,req}}} & \text{if } N^{\text{hosp,req}} > C^{\text{hosp}} - O^{\text{hosp}}
\end{cases}$$

where:

- $N^{\text{hosp,req}}$ is the total number requiring treatment
- $C^{\text{hosp}}$ is the hospital bed capacity
- $O^{\text{hosp}}$ is the current hospital bed occupancy

## ICU Bed Availability

The probability of receiving mechanical ventilation is:

$$p^{\text{ICU}} = \begin{cases}
1 & \text{if } N^{\text{ICU,req}} \leq C^{\text{ICU}} - O^{\text{ICU}} \\
\frac{C^{\text{ICU}} - O^{\text{ICU}}}{N^{\text{ICU,req}}} & \text{if } N^{\text{ICU,req}} > C^{\text{ICU}} - O^{\text{ICU}}
\end{cases}$$

where:

- $N^{\text{ICU,req}}$ is the total number requiring mechanical ventilation
- $C^{\text{ICU}}$ is the ICU bed capacity
- $O^{\text{ICU}}$ is the current ICU bed occupancy

# Time-Varying Parameters

Several parameters in the model are time-varying and implemented using interpolation:

- $\beta(t)$: Transmission rate
- $\gamma_R(t)$: Rate of immunity waning
- $\xi^{\text{inf}}_v(t)$: Vaccine efficacy against infection
- $\xi^{\text{dis}}_v(t)$: Vaccine efficacy against disease
- $\gamma^{\text{vac}}_v(t)$: Vaccine waning rates
- $\gamma^{\text{Hosp,surv}}(t)$, $\gamma^{\text{Hosp,die}}(t)$: (non-ICU) Treatment progression rates
- $\gamma^{\text{ICU,surv}}(t)$, $\gamma^{\text{ICU,die}}(t)$: ICU progression rates
- $m^{\text{hosp}}(t)$: Hospitalization probability multiplier
- $m^{\text{sev}}(t)$: Severity probability multiplier

# Parameter Definitions

The following table summarizes the key model parameters:

```{r parameter-table, echo=FALSE}
library(knitr)

parameter_table <- data.frame(
  Parameter = c(
    "$\\gamma_E$",
    "$\\gamma^{\\text{Mild}}$",
    "$\\gamma^{\\text{Case}}$",
    "$\\gamma^{\\text{rec}}$",
    "$p^{\\text{hosp}}_a$",
    "$p^{\\text{sev}}_a$",
    "$p^{\\text{death,Hosp,treat}}_a$",
    "$p^{\\text{death,Hosp,notreat}}_a$",
    "$p^{\\text{death,ICU,treat}}_a$",
    "$p^{\\text{death,ICU,notreat}}_a$"
  ),
  Description = c(
    "Rate of progression through latent infection",
    "Recovery rate from mild infection",
    "Rate of progression from severe infection",
    "Rate of progression through post-ICU recovery",
    "Probability of hospitalization by age group",
    "Probability of severe disease (requiring ICU treatment) by age group",
    "Death probability with (non-ICU) treatment",
    "Death probability without ICU treatment",
    "Death probability with mechanical ventilation",
    "Death probability without mechanical ventilation"
  )
)

kable(parameter_table, 
      caption = "Key model parameters",
      col.names = c("Parameter", "Description"),
      escape = FALSE)
```

# Acknowledgements

This document was initially generated using Claude and Copilot, then edited for clarity and correctness.
